### 2021-01-22 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Done for the day!

Ideally, we would have simultaneously observed the same targets with \ktwo\ and \tess\ to measure the instantaneous amplitudes for individual stars.  Cycle 4 provides a unique opportunity to compare the amplitudes of \tess\ and \ktwo\ for the largest sample to date.

The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.  Other important-albeit-unknown factors like stellar inclination angle and multiplicity \emph{cancel out} when we compare the same stars head-to-head.  We, therefore, propose to follow \cite{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.  The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in Figure \ref{fig:expected}.

\tess\ Cycle 4 will observe XX sources in the \ktwo\ campaign fields.  The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles. On average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.

\section{Technical Feasibility: Consistency is key}

Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.

The yield of spotted stars in \tess\ will number in the hundreds of thousands.  The \tess\ and \ktwo\ missions both target cool main sequence stars with an abundance of K and M dwarfs.  This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in panel D of Figure \ref{fig:predictAmps}.  This program will attempt---on a best effort basis---to measure slower rotating stars with ($P_{rot}<20$ days), which have periods comparable to typical observation windows of 27$-$54 days, and amplitudes comparable to the precision limits of \tess.

@@

The main deliverable of this work will be a probabilistic assessment of the starspot temperature and starspot coverage faction for every star with a measured starspot modulation period and amplitude in \tess.  These probability distributions will empower exoplanet researchers to correct for biases in exoplanet transit radii.  The probability distributions will reflect all the available information garnered from \tess\ and reasonable assumptions about starspot geometrical distributions.

The starspot properties are also of intrinsic astrophysical interest.  Trends in $T_{spot}$ with rotation rate and mass can reveal important clues to stellar activity, magnetic fields, and stellar atmospheres.  This work will directly address starspot-induced scatter and bias in isochronal-derived stellar ages \cite{2015ApJ...807..174S, 2017ApJ...836..200G}.

@@

PI will carry out the data analysis using custom scripts that build upon the heritage of the \ktwo\ data analysis ecosystem, including \texttt{lightkurve} and \texttt{celerite}.  The PI will first apply these scripts to \ktwo\ lightcurves.  Undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.  Half of the funding will go to 1 undergraduate and 1 graduate student Summer 2020 salary, and 0.3 FTE to support the PI.  All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
PI will provide students with an interactive tool to visualize periodic lightcurve structures to aid rapid visual spot-checking of the multidimensional analysis output.  Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature and coverage fraction, 5) addition model fitting metadata.  PI will co-advise---with Co-I---students in the writeup of the paper(s), authored and/or coauthored by students.

>>>>

The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.  Other important-albeit-unknown factors like stellar inclination angle and multiplicity \emph{cancel out} when we compare the same stars head-to-head.  We, therefore, construct a plot of amplitude versus period, grouped by spectral type bins.  The \tess\ and \ktwo\ samples should look identical in these plots, but the locus of \tess\ sources resting systematically below the \ktwo\ sample, as in Figure \ref{fig:expected}.

The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles, altering their amplitudes in 2021 compared to say 2018. So it is therefore impossible to assign a starspot contrast ratio $\mathcal{R}$ for any individual star.  Instead, we propose to measure the mean value of \emph{ensembles} of thousands of stars, again grouped by sources of similar properties.  On average stars will enter and exit stellar activity cycles in equal measure injecting per-star variance, but keeping the subsample mean $\mathcal{R}$ the same.

\section{Technical Feasibility: Consistent and Systematic}

Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.  Amplitude estimates exist for some clusters, but with differing heuristics for for estimating the variability, such as interquartile range, sine-wave fitting, kernel methods, peak-to-valley, Fourier-based approaches, and other techniques that cannot be robustly intercompared to new \tess\ data.  One centerpiece for this project is to compute modulation amplitudes systematically across both \ktwo\ and \tess.

The \ktwo\ mission targeted cool main sequence stars with an abundance of K and M dwarfs.  This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in Figure \ref{fig:expected}.  We estimate the total number of conceivable \tess\ and \ktwo\ rotationally modulated stars to be of order $10^5$, including cluster members, flare stars, active M-dwarfs, bright planet candidates, and other community targeted stellar programs acquired with long cadence postage stamps in \ktwo.  This parent sample has to be vetted for \tess\ background overcrowding, \tess\ scattered light, variability classification provenance, signal-to-noise ratio, and other quality assurance checks. The mere curation of this sample represents a portion of the proposed work plan.

@@

The main deliverable of this work will be a probabilistic assessment of the starspot contrast ratio between \kepler\ and \tess\ as a function of spectral type and rotation rate. These probability distributions will empower exoplanet researchers to quantify biases in exoplanet transit radii, answering the question ``What starspot emission properties are plausible for my host-star of interest?''.  For example, this work will make it possible to dispel certain categories of spot temperatures that could otherwise imbue spurious features in a JWST spectrum.

The starspot properties are also of intrinsic astrophysical interest.  Trends in $T_{spot}$ with rotation rate and mass can reveal important clues to stellar activity, magnetic fields, and stellar atmospheres.  This work will inform starspot-induced scatter and bias in isochronal-derived stellar ages \cite{2015ApJ...807..174S, 2017ApJ...836..200G}, since spotted star SEDs can be tuned to fit the trends derived here.

@@

PI will carry out the data analysis using a combination of new and existing scripts built upon the heritage of the \ktwo\ data analysis ecosystem, including \texttt{lightkurve} and \texttt{celerite}.  The PI will first apply scripts to both \ktwo\ and \tess\ lightcurves.  Undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.  Half of the funding will go to 1 undergraduate and 1 graduate student Summer 2020 salary, and 0.3 FTE to support the PI.  All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
PI will provide students with an interactive tool to visualize periodic lightcurve structures to aid rapid visual spot-checking of analysis output.  Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature spot contrast ratio, 5) addition model fitting metadata.  PI will co-advise---with Co-I---students in the writeup of the paper(s), authored and/or coauthored by students.
